---
title: "Untitled"
author: "Richard D. Morey"
date: "14/11/2016"
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

df = as.integer( (nrow(m) - 1L) * (ncol(m) - 1L) )

contcor = contcor & (df==1)  
```


```{r echo=FALSE, warning=FALSE}
z = chisq.test(m, correct = contcor)
```

### Our matrix

```{r echo=FALSE}
knitr::kable(m, format = "html", table.attr = "class='ct_test_table'")
```

### Computing the expected entries

```{r echo=FALSE}
knitr::kable(z$expected, format = "html", table.attr = "class='ct_test_table'")
```

```{r child='child_continuity.Rmd', eval = contcor}
```

```{r child='child_no_continuity.Rmd', eval = !contcor}
```

### Standardizing the squared deviations


```{r echo=FALSE}
knitr::kable(res^2/z$expected, format = "html", table.attr = "class='ct_test_table'")

res2 = res^2/z$expected

X2 = sum(res^2/z$expected)
```

### The *X*<sup>2</sup> statistic

The sum of the squared, standardized deviations from the expected values is
\[
\begin{eqnarray}
X^2 &=& `r paste(round(res2,3), collapse = "+")`\\
&=& `r round(X2, 3)`
\end{eqnarray}
\]

Because the contingency table has `r nrow(m)` rows and `r ncol(m)` columns, the $X^2$ statistic has $(`r nrow(m)` - 1)\times (`r ncol(m)` - 1) = `r df`$ degrees of freedom. We report this as $X^2(`r df`) = `r round(X2, 3)`$.




