---
title: "Untitled"
author: "Richard D. Morey"
date: "14/11/2016"
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

df = as.integer( (nrow(m) - 1L) * (ncol(m) - 1L) )

contcor = contcor & (df==1)  
```


```{r echo=FALSE, warning=FALSE}
z = chisq.test(m, correct = contcor)
```

### Our matrix

```{r echo=FALSE}
knitr::kable(m, format = "html", table.attr = "class='ct_test_table'")
```

### Computing the expected entries

In order to compute the expected entries, we need to estimate that marginal probabilities for the rows and columns. First we note that there are a total of $n = `r sum(m)`$ observations across all cells.

#### Marginal row probabilities

To find the marginal row probabilities, we sum the observations across all columns for each row. For instance, the sum for the first row is
\[
`r paste(m[1,],collapse=" + ")` = `r rowSums(m)[1]`
\]

The proportion of all observations that are in the first row is $`r rowSums(m)[1]`/`r sum(m)` = `r round(rowSums(m)[1]/sum(m),3)`$. The proportion of observations across all rows is
```{r echo=FALSE}
rows = rbind(
  paste("$r_{",1:nrow(m),"}$"),
  paste("\\[ \\frac{",rowSums(m),"}{",sum(m),"}=",
        round(rowSums(m)/sum(m),3),"\\]")
  )

knitr::kable(rows, format = "html", table.attr = "class='ct_test_table'")
```

#### Marginal column probabilities

Performing the same calculations for the columns as we did for the rows, we obtain the column proportions
```{r echo=FALSE}
cols = rbind(
  paste("$c_{",1:ncol(m),"}$"),
  paste("\\[ \\frac{",colSums(m),"}{",sum(m),"}=",
        round(colSums(m)/sum(m),3),"\\]")
  )

knitr::kable(cols, format = "html", table.attr = "class='ct_test_table'")
```

#### Assuming independence

Suppose that the proportions that we computed above were the true marginal proportions of observations in each row or column. How many observations would we predict in the cell in the first row, first column? Using the multiplication rule for independent probabilities, we would compute the true proportion in row 1, column 1 as
\[
r_1 \times c_1 = `r round(rowSums(m)[1]/sum(m),3)` \times `r round(colSums(m)[1]/sum(m),3)` = `r round(rowSums(m)[1]/sum(m) * colSums(m)[1]/sum(m), 3)` 
\]

That is we would expect `r round(100*rowSums(m)[1]/sum(m) * colSums(m)[1]/sum(m),1)`% of observations to be in the cell in the first row, first column. Given that there are `r sum(m)` total observations, we would expect
\[
`r round(rowSums(m)[1]/sum(m) * colSums(m)[1]/sum(m),3)` \times `r sum(m)` = `r round(z$expected[1,1],3)`
\]
observations in the first row, first column.

We can repeat this computation for every row and column and create an "expected" count in each cell:

```{r echo=FALSE}
knitr::kable(z$expected, format = "html", table.attr = "class='ct_test_table'")
```


```{r child='child_continuity.Rmd', eval = contcor}
```

```{r child='child_no_continuity.Rmd', eval = !contcor}
```

### Standardizing the squared deviations


```{r echo=FALSE}
knitr::kable(res^2/z$expected, format = "html", table.attr = "class='ct_test_table'")

res2 = res^2/z$expected

X2 = sum(res^2/z$expected)
```

### The *X*<sup>2</sup> statistic

The sum of the squared, standardized deviations from the expected values is
\[
\begin{eqnarray}
X^2 &=& `r paste(round(res2,3), collapse = "+")`\\
&=& `r round(X2, 3)`
\end{eqnarray}
\]

Because the contingency table has `r nrow(m)` rows and `r ncol(m)` columns, the $X^2$ statistic has $(`r nrow(m)` - 1)\times (`r ncol(m)` - 1) = `r df`$ degrees of freedom. We report this as $X^2(`r df`) = `r round(X2, 3)`$.




