
### $\phi$ correlation coefficient

The $\phi$ correlation coefficient is a measure of the tendency of the observations to be either on the diagonal of the $2\times2$ contingency table (positive correlation) or on the off diagonal (negative correlation).

\[
\frac{x_{11}x_{22} - x_{12}x_{21}}{\sqrt{n_{\cdot1}n_{\cdot2}n_{1\cdot}n_{2\cdot}}}
\]

where $n_{1\cdot}$ means the number of observations in the first row, $n_{2\cdot}$ the number observations in the second row, and $n_{\cdot1}$ and $n_{\cdot2}$ are the numbers of observations in the first and second column, respectively.

For our matrix, this is:

\[
\frac{`r m[1,1]` \times `r m[2,2]` - `r m[1,2]`\times `r m[2,1]`}{\sqrt{`r colSums(m)[1]`
 \times `r colSums(m)[2]`
 \times `r rowSums(m)[1]`
 \times `r rowSums(m)[2]`
}} = 
\frac{`r m[1,1]*m[2,2] - m[1,2]*m[2,1]`}{`r round(sqrt( colSums(m)[1]*colSums(m)[2]*rowSums(m)[1]*rowSums(m)[2]), 3)`
} = `r round((m[1,1]*m[2,2] - m[1,2]*m[2,1])/sqrt( colSums(m)[1]*colSums(m)[2]*rowSums(m)[1]*rowSums(m)[2]), 3)`
\]


The $\phi$ correlation coefficient was computed using the `phi` function in the `psych` R package ([CRAN](https://cran.r-project.org/web/packages/psych)).

### Yule's $Q$ correlation coefficient

Yule's $Q$ correlation coefficient was computed using the `Yule` function in the `psych` R package ([CRAN](https://cran.r-project.org/web/packages/psych)).

### Tetrachoric correlation coefficient

The tetrachoric correlation is computed assuming that the dichotomous data in the 2x2 table arise from a latent bivariate normal with a mean vector $(0,0)$, marginal variances of 1, and a correlation coefficient of $\rho$. Call this bivariate observation $(z_1, z_2)$.

The discrete observations are assumed to arise from these latent bivariate normal observations by means of two criteria: one for the columns and one for the rows, denoted $\tau_1$ and $\tau_2$ respectively. If $z_1\leq\tau_1$, then the observation will appear in the first row; otherwise it will appear in the second row. If $z_2\leq\tau_2$, then the observation will appear in the first column; otherwise it will appear in the second column. 

The parameters $\rho$, $\tau_1$ and $\tau_2$ are selected to fit the observed joint probabilities in the contingency table, which are:

```{r echo=FALSE}
m.latex = paste("\\[",
                paste(
                  paste("\\frac{",as.character(m),"}{",sum(m),"}="),
                  paste(round(m/sum(m),3))
                ),
                "\\]")
dim(m.latex) = c(2,2)
knitr::kable(m.latex, format = "html", table.attr = "class='ct_test_table'")
```

```{r echo=FALSE}
tc = psych::tetrachoric(m)

```


```{r}
xx <- seq(qnorm(.001), qnorm(.999), len = 100)
S = matrix(c(1,tc$rho,tc$rho,1),2,2)
M = c(0,0)
yy <- outer(xx, xx, Vectorize(function(x,y){
  mvtnorm::dmvnorm( c(x,y), M, S)  
},c("x","y")))
```


```{r include = FALSE}
crits.rows = c(-Inf,tc$tau[1],Inf)
crits.cols = c(-Inf,tc$tau[2],Inf)
b = expand.grid(row = 1:nrow(m),col = 1:ncol(m))
b$lower.rows = crits.rows[b$row]
b$upper.rows = crits.rows[b$row+1]
b$lower.cols = crits.cols[b$col]
b$upper.cols = crits.cols[b$col+1]
probs = apply(b,1,function(v){
  lower = c(v['lower.rows'], v['lower.cols'])
  upper = c(v['upper.rows'], v['upper.cols'])
  mvtnorm::pmvnorm(lower = lower, upper = upper,mean = M, sigma = S)
})

dim(probs) = dim(m)

```

Here are the best fitting joint probabilities:

```{r echo=FALSE}
knitr::kable(round(probs,3), format = "html", table.attr = "class='ct_test_table'")
```


The plot below shows a representation of the best fitting tetrachoric correlation model. It represents a bivariate normal distribution with a correlation coefficient of $\rho=`r round(tc$rho,3)`$.

```{r fig.width = 5, fig.height=5}
contour(xx,xx,yy, asp=TRUE, ylab=expression(paste(z[2], " (columns)")),
        xlab=expression(paste(z[1], " (rows)")))
abline(v = tc$tau[1], col="red")
text(tc$tau[1], par()$usr[3], expression(tau[1]), srt=90, adj = c(-.2,-.2), col="red")

abline(h = tc$tau[2], col="red")
text(par()$usr[1], tc$tau[2], expression(tau[2]), adj = c(-.2,-.2), col="red")
text(par()$usr[1], par()$usr[3],expression(x[11]), adj=c(-0.2,-0.2),cex=2, col=rgb(0,0,0,.5))
text(par()$usr[2], par()$usr[4],expression(x[22]), adj=c(1.2,1.2),cex=2, col=rgb(0,0,0,.5))
text(par()$usr[2], par()$usr[3],expression(x[21]), adj=c(1.2,-.2),cex=2, col=rgb(0,0,0,.5))
text(par()$usr[1], par()$usr[4],expression(x[12]), adj=c(-.2,1.2),cex=2, col=rgb(0,0,0,.5))

rect(tc$tau[1],par()$usr[3],par()$usr[2],par()$usr[4],col=rgb(0,0,0,.05), border = NA)

```

Notice that the plot is rotated relative to the matrix, because in plots we put the origin in the lower left corner, while in contingency tables the "origin" is in the upper left. The rows are shaded as in the matrix to make this clearer.

The tetrachoric correlation coefficient was computed using the `tetrachoric` function in the `psych` R package ([CRAN](https://cran.r-project.org/web/packages/psych)).
