
```{r echo=FALSE}
require(PropCIs)
```

```{r echo=FALSE}
or.est = function(x1, n1, x2, n2, ...){
  or = (x1 / (n1 - x1) ) / ( x2 / (n2 - x2) )
  se = sqrt( 1/x1 + 1/n1 + 1/x2 + 1/n2 )
  list(or = or, se = se)
}

rr.est = function(x1, n1, x2, n2, ...){
  rr = ( x1 / n1 ) / ( x2 / n2 )
  se = sqrt( 1/x1 + 1/n1 + 1/x2 + 1/n2 )
  list(rr = rr, se = se)
}

dp.est = function(x1, n1, x2, n2, add = c(0,0), ...){
  phatx = ( (x1+add[1]) / (n1 + sum(add)) )
  phaty = ( (x2+add[1]) / (n2 + sum(add)) )
  dp = phatx - phaty
  se = sqrt(phatx * (1-phatx) / (n1 + sum(add)) + phaty * (1-phaty) / (n2 + sum(add)))
  list(dp = dp, se = se)
}

```

### Difference in proportions

```{r echo=FALSE}
DP = do.call(dp.est, ci.args)
```

If $p_d = p_1 - p_2$, we can estimate $p_d$ with
\[
\begin{eqnarray}
\hat{p_d} &=& \frac{x_1}{n_1}-\frac{x_2}{n_2}\\
&=& \frac{`r ci.args$x1`}{`r ci.args$n1`}-\frac{`r ci.args$x2`}{`r ci.args$n2`}\\
&=& `r round(DP$dp, 3)`
\end{eqnarray}
\]



#### Wald interval on difference

The Wald interval uses a normal approximation to the binomial distribution. With a large enough sample and probabilities that are not too extreme,
\[
\frac{x_1}{n_1}\stackrel{\cdot}{\sim} \mbox{Normal}\left(p_1, \sqrt\frac{\hat{p}_1(1-\hat{p}_1)}{n_1}\right),
\]
and likewise for $p_2$. Assuming that the observations are independent, an estimate of the standard error of $x_1/n_1 - x_2/n_2$ is
\[
\sqrt{\frac{\hat{p}_1(1-\hat{p}_1)}{n_1} + \frac{\hat{p}_2(1-\hat{p}_2)}{n_2}}
\]



```{r echo=FALSE}
ci.args2 = ci.args
ci.args2$adjust = "Wald"
ci = do.call(PropCIs::wald2ci, ci.args2)
```

The Wald `r 100*confcoef`% confidence interval is 
\[
\begin{eqnarray}
`r round(DP$dp, 3)`\pm `r round(zstar,3)`\times`r round(DP$se,3)`\\
[`r round(DP$dp - zstar*DP$se,3)`,`r round(DP$dp + zstar*DP$se,3)`]
\end{eqnarray}
\]


```{r echo=FALSE}
ci.args2 = ci.args
ci.args2$adjust = "AC"
ci.args2$add = c(1,1)
DP2 = do.call(dp.est, ci.args2)
ci.args2$add = NULL
ci = do.call(PropCIs::wald2ci, ci.args2)
```

With Agresti and Caffo (2000) adjustment (adding one failure and one success). With this adjustment we estimate $p_d$ with
\[
\begin{eqnarray}
\hat{p_d} &=& \frac{x_1+1}{n_1+2}-\frac{x_2+1}{n_2+2}\\
&=& \frac{`r ci.args$x1+1`}{`r ci.args$n1+2`}-\frac{`r ci.args$x2+1`}{`r ci.args$n2+2`}\\
&=& `r round(DP2$dp, 3)`
\end{eqnarray}
\]
The standard error is similarly estimated based on the adjusted counts.

The adjusted Wald `r 100*confcoef`% confidence interval is 
\[
\begin{eqnarray}
`r round(DP2$dp, 3)`\pm `r round(zstar,3)`\times`r round(DP2$se,3)`\\
[`r round(DP2$dp - zstar*DP2$se,3)`,`r round(DP2$dp + zstar*DP2$se,3)`]
\end{eqnarray}
\]


The Wald and adjusted Wald intervals were computed using the `wald2ci` function in the `PropCIs` R package ([CRAN](https://cran.r-project.org/web/packages/PropCIs)).



#### Score interval on difference

```{r echo=FALSE}
do.call(PropCIs::diffscoreci, ci.args)
```

This score interval was computed using the `diffscoreci` function in the `PropCIs` R package ([CRAN](https://cran.r-project.org/web/packages/PropCIs)).



### Relative risk

```{r echo=FALSE}
RR = do.call(rr.est, ci.args)
```
\[
\begin{eqnarray}
\hat{RR} &=& \left.\frac{x_1}{n_1}\middle/\frac{x_2}{n_2}\right.\\
&=& \left.\frac{`r ci.args$x1`}{`r ci.args$n1`}\middle/\frac{`r ci.args$x2`}{`r ci.args$n2`}\right.\\
&=& `r round(RR$rr, 3)`
\end{eqnarray}
\]



#### Score interval

```{r echo=FALSE}
do.call(PropCIs::riskscoreci, ci.args)
```

This score interval was computed using the `riskscoreci` function in the `PropCIs` R package ([CRAN](https://cran.r-project.org/web/packages/PropCIs)).


### Odds ratio

```{r echo=FALSE}
OR = do.call(or.est, ci.args)
```
\[
\begin{eqnarray}
\hat{OR} &=& \left.\frac{x_1}{n_1 - x_1}\middle/\frac{x_2}{n_2 - x_2}\right.\\
&=& \left.\frac{`r ci.args$x1`}{`r ci.args$n1 - ci.args$x1`}\middle/\frac{`r ci.args$x2`}{`r ci.args$n2 - ci.args$x2`}\right.\\
&=& `r round(OR$or, 3)`
\end{eqnarray}
\]

#### Score interval


```{r echo=FALSE}
do.call(PropCIs::orscoreci, ci.args)
```

This score interval was computed using the `orscoreci` function in the `PropCIs` R package ([CRAN](https://cran.r-project.org/web/packages/PropCIs)).

