
```{r echo=FALSE}
require(PropCIs)
```

```{r echo=FALSE}
or.est = function(x1, n1, x2, n2, ...){
  or = (x1 / (n1 - x1) ) / ( x2 / (n2 - x2) )
  se = sqrt( 1/x1 + 1/n1 + 1/x2 + 1/n2 )
  list(or = or, se = se)
}
```

### Difference in proportions


#### Wald interval on difference

```{r}
ci.args2 = ci.args
ci.args2$adjust = "Wald"
do.call(PropCIs::wald2ci, ci.args2)
```

With Agresti and Caffo (2000) correction (adding one failure and one success):

```{r}
ci.args2 = ci.args
ci.args2$adjust = "AC"
do.call(PropCIs::wald2ci, ci.args2)
```

#### Score interval on difference

```{r}
do.call(PropCIs::diffscoreci, ci.args)
```


### Relative risk

#### Score interval

```{r}
do.call(PropCIs::riskscoreci, ci.args)
```


### Odds ratio

```{r}
OR = do.call(or.est, ci.args)
```
\[
\begin{eqnarray}
\hat{OR} &=& \left.\frac{x_1}{n_1 - x_1}\middle/\frac{x_2}{n_2 - x_2}\right.\\
&=& \left.\frac{`r ci.args$x1`}{`r ci.args$n1 - ci.args$x1`}\middle/\frac{`r ci.args$x2`}{`r ci.args$n2 - ci.args$x2`}\right.\\
&=& `r round(OR$or, 3)`
\end{eqnarray}
\]

#### Score interval


```{r}
do.call(PropCIs::orscoreci, ci.args)
```

This score interval was computed using the `orscoreci` function in the `PropCIs` R package ([CRAN](https://cran.r-project.org/web/packages/PropCIs)).

